---
title: "Exploratory Data Analysis of Global Data from Gampminder in R"
output: html_document
author: "Hari Cheboyina"
---
###Gapminder Package Description:
* 1704 observations
* 6 variables
    + country: a factor with 142 levels
    + continent: a factor with 5 levels
    + year: going from 1952 to 2007 in increments of 5 years
    + pop: population
    + gdpPercap: GDP per capita
    + lifeExp: life expectancy  

###Libraries Used:
* Gapminder
* ggplot2
* dplyr
    
***
> Import libraries

```{r, message = FALSE}
library(gapminder)
library(ggplot2)
library(dplyr)


```
##### Summary of Gapminder Data 
```{r, message = FALSE}
summary(gapminder)
```
summary(dataframe) results in Minimum, 1st Quartile, Median, Mean, 3rd Quartile, Maximum of the dataframe.

> Comparing Variables

##### Now that we have imported the data, lets compare two variables. 
###### I wonder how life expectancy varies in the two most populus countries in the world. Lets compare the life expectancy rate of China and India side by side using ggplot
```{r, message = FALSE}
gapminder %>%
  select(country, lifeExp, year, pop) %>%
  filter(country =="China" |
        country == "India") %>%
ggplot(aes(x=country, y=lifeExp, col=year, size=log(pop)))+
  geom_point(alpha=0.4)+
  geom_smooth(method=lm)
```
Well we can observe that the life expectancy has increased in both the countries over the years. We can observe that the people in china live longer than people in India. Well, this is trickier thing to conclude based on this graph, but if you google the infant mortality rate in India, you will know why. (Unless you understand the relation between infant mortality rate and life expectancy)

***
####How about relationship between gdp(Gross domestic product) and life expectancy? Does more GDP of a country indicate it's people can live longer and viceversa? Lets find out.

Lets filter the countries with GDP lessthan $50000 and plot a graph using ggplot and wrap the countries within their respective continents.
```{r}
gapminder %>%
  filter(gdpPercap<50000) %>%
  ggplot(aes(x=log(gdpPercap), y=lifeExp, col=year, size=pop))+
  geom_point(alpha=0.3)+
  geom_smooth(method=lm)+
  facet_wrap(~continent)
```

```{r}
gapminder %>%
  filter(year == 2007) %>%
  group_by(continent) %>%
  summarise(lifeExp = median(lifeExp))
```

```{r}
ggplot(gapminder, aes(x = continent, y = lifeExp)) +
  geom_boxplot(outlier.colour = "hotpink") +
  geom_jitter(position = position_jitter(width = 0.1, height = 0), alpha = 1/4)
```

```{r}
gap_with_colors <-
  data.frame(gapminder,
             cc = I(country_colors[match(gapminder$country,
                                         names(country_colors))]))

# bubble plot, focus just on Africa and Europe in 2007
keepers <- with(gap_with_colors,
                continent %in% c("Asia", "Europe") & year == 2007)
plot(lifeExp ~ gdpPercap, gap_with_colors,
     subset = keepers, log = "x", pch = 21,
     cex = sqrt(gap_with_colors$pop[keepers]/pi)/1500,
     bg = gap_with_colors$cc[keepers])

```

```{r}
ggplot(subset(gapminder, continent != "Americas"))+
       aes(x = year, y = lifeExp, group = country, color = country) +
  geom_line(lwd = 1, show.legend = FALSE) + facet_wrap(~ continent) +
  scale_color_manual(values = country_colors) +
  theme_bw() + theme(strip.text = element_text(size = rel(1.1)))
```